---
title: "Happy Little Clouds Example"
author: "Lucy D'Agostino McGowan"
date: "2019-08-29"
output: html_document
---

In this little example we use data from the FiveThirtyEight article [A Statistical Analysis of the Work of Bob Ross](https://fivethirtyeight.com/features/a-statistical-analysis-of-the-work-of-bob-ross/). 

## Load data and packages

```{r load-packages, message = FALSE}
library(fivethirtyeight)
library(tidyverse)
```

This dataset contains information about `r nrow(bob_ross)` paintings from Bob Ross's television series painted over `r bob_ross %>% summarise(max = max(season)) %>% pull()` seasons.

Let's see how many episodes there were in each season.

```{r}
bob_ross %>%
  count(season)
```

There are --- episodes in each season.

The variables we are going to focus on are:

- `cloud`: whether the painting contained a happy little cloud
- `ocean`: whether the painting was of the ocean
- `season`: the season the painting was created.

## Analysis

Let's sum up the number of cloud paintings in each season.

```{r}
bob_ross %>%
  group_by(season) %>%
  summarise(n_clouds = sum(clouds))
```

Now let's tally the number of the ocean paintings by season.

```{r}
bob_ross %>%
  group_by(season) #%>%
 # summarise(---)
```


How often were there ocean paintings with clouds?

Let's create a new variable called `ocean_and_clouds`.

```{r}
bob_ross <- bob_ross %>%
  mutate(ocean_and_clouds = case_when(
    ocean == 1 & clouds == 1 ~ TRUE,
    TRUE ~ FALSE)
  )
```

```{r}
bob_ross %>%
  group_by(season) %>%
  summarise(n_ocean_and_clouds = sum(ocean_and_clouds))
```

What percent of ocean paintings have clouds?

```{r}
bob_ross %>%
  summarise(n_ocean_and_clouds = sum(ocean_and_clouds),
            n_ocean = sum(ocean),
            pct = n_ocean_and_clouds / n_ocean)
```

What percent of cloud paintings have an ocean?

```{r}
# bob_ross %>%
  # summarise(---)
```

Let's visualize the number of paintings with happy little clouds over time.

```{r}
summarise_bob_ross <- bob_ross %>%
  group_by(season) %>%
  summarise(clouds = sum(clouds))
```

```{r}
ggplot(data = summarise_bob_ross, mapping = aes(season, clouds)) + 
  geom_point() + 
  labs(title = "Happy little cloud paintings by season",
       x = "Season",
       y = "---")
```

```{r}
ggplot(data = summarise_bob_ross, mapping = aes(season, clouds)) + 
  geom_point() + 
  labs(title = "Happy little cloud paintings by season",
       subtitle = "---",
       x = "Season",
       y = "# of cloud paintings") + 
  geom_smooth(method = "loess")
```

