---
title: "Welcome to Statistical Models"
author: "Dr. D'Agostino McGowan"
date: "2019-08-20"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/icon.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r child = "setup.Rmd"}
```

layout: true
  
<div class="my-footer">
<span>
Slides adpated from <a href="https://github.com/Sta199-S18/website" target="_blank">Dr. Mine Çetinkaya-Rundel</a> by Dr. Lucy D'Agostino McGowan
</span>
</div> 

---

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Who am I?

Lucy D'Agostino McGowan

<i class="fa fa-envelope"></i> &nbsp; [mcgowald@wfu.edu](mailto:mcgowald@wfu.edu) <br>
<i class="fa fa-home"></i> &nbsp; [bit.ly/sta-212-f19](https://sta-212-f19.netlify.com) <br>
<i class="fa fa-university"></i> &nbsp; Tribble A13 <br>
<i class="fa fa-calendar"></i> &nbsp; Wed 2:00-3:00p

---

# What is this course?

Everything you want to know about the course, and everything you will need for the course will be posted at

[bit.ly/sta-212-f19](http://bit.ly/sta-212-f19)

--

- Will we be doing computing? Yes.

--

- We expect that you have taken at least 1 Stat course (either at Wake Forest or AP Statistics)

--

- What computing language will we learn? R.

--

- Why not language X? We can discuss that over `r emo::ji("coffee")`.


---

class: center, middle

# Data in the wild

---

# A year as told by fitbit

by Nick Strayer

http://livefreeordichotomize.com/2017/12/27/a-year-as-told-by-fitbit/

---

# R-Ladies global tour

by Maelle Salmon

http://www.masalmon.eu/2017/10/06/globalrladiestour/

---

# Text analysis of Trump's tweets confirms he writes only the (angrier) Android half

by David Robinson (Stack Overflow)

http://varianceexplained.org/r/trump-tweets/

---

class: center, middle

# Your turn!

---

## Create a GitHub account

Go to https://github.com/, and create an account (unless you already have one). Tips for selecting a username:<sup>1</sup>

- Incorporate your actual name! People like to know who they’re dealing with. Also makes your username easier for people to guess or remember.
- Reuse your username from other contexts if you can, e.g., Twitter or Slack.
- Pick a username you will be comfortable revealing to your future boss.
- Shorter is better than longer.
- Be as unique as possible in as few characters as possible. In some settings GitHub auto-completes or suggests usernames.
- Make it timeless. Don’t highlight your current university, employer, or place of residence.
- Avoid words laden with special meaning in programming, like `NA`.

.footnote[ [1] Source: [Happy git with R](http://happygitwithr.com/github-acct.html#username-advice) by Jenny Bryan.]

<font color="#E34132">
Once done, place a green sticky on your laptop. If you have questions, place a pink sticky.
</font>

---

## Join RStudio.cloud

Go to [bit.ly/sta-212-f19-rstudio-join](http://bit.ly/sta-212-f19-rstudio-join), and log in with your GitHub credentials.

<font color="#E34132">
Once done, place a green sticky on your laptop. If you have questions, place a pink sticky.
</font>
---

## Create your first data visualization

- Once you log on to RStudio Cloud, click on this course's workspace "STA 199 - Spring 18"
- You should see a project called UN Votes, fork it by clicking on the <i class="fa fa-fork"></i> icon. This will create your copy of the project and launch it.
- In the Files pane in the bottom right corner, spot the file called `unvotes.Rmd`. Open it, and then click on the "Knit" button.
- Go back to the file and change your name on top (in the `yaml` -- we'll talk about what this means later) and knit again.
- Then change the country names to those you're interested in. Your spelling and capitalization should match the data so take a peek at the Appendix to see how the country names are spelled. Knit again. And voila, your first data visualization!

<font color="#E34132">
Once done, place a green sticky on your laptop. If you have questions, place a pink sticky.
</font>
---

class: center, middle

# Course structure and policies

---

## Class meetings

- Interactive

- Some lectures, lots of learn-by-doing

- Bring your laptop to class every day

---

## Diversity & Inclusiveness:

- Intent: Students from all diverse backgrounds and perspectives be well-served by this course, that students' learning needs be addressed both in and out of class, and that the diversity that the students bring to this class be viewed as a resource, strength and benefit. It is my intent to present materials and activities that are respectful of diversity: gender identity, sexuality, disability, age, socioeconomic status, ethnicity, race, nationality, religion, and culture. Let me know ways to improve the effectiveness of the course for you personally, or for other students or student groups.

- If you have a name and/or set of pronouns that differ from those that appear in your official Wake Forest records, please let me know!

- If you feel like your performance in the class is being impacted by your experiences outside of class, please don't hesitate to come and talk with me. I want to be a resource for you. If you prefer to speak with someone outside of the course, your academic dean is an excellent resource. 

- I (like many people) am still in the process of learning about diverse perspectives and identities. If something was said in class (by anyone) that made you feel uncomfortable, please talk to me about it.

---

## How to get help

- Course content, logistics, etc. discussion via GitHub on the [sta-212-f19/community](https://github.com/sta-212-f19/community) repository. Note that this is a public discussion forum, which means others outside of the course can stumble upon it and help you as well.
    - See course policies for tips on posting questions.

- For personal and grade related questions, use email.

---

## Academic integrity

Adhere to the Wake Forest Honor Code. Academic dishonesty will not be tolerated.

---

## Sharing/reusing code

* There are many online resources for sharing code (for example, StackOverflow) - you **may** use these resources but **must explicitly cite** where you have obtained code (both code you used directly and "paraphrased" code / code used as inspiration). Any reused code that is not explicitly cited will be treated as plagiarism.
* You **may** discuss the content of assignments with others in this class. If you do so, please acknowledge your collaborator(s) at the top of your assignment, for example: "Collaborators: Gertrude Cox, Florence Nightingale David". Failure to acknowledge collaborators will result in a grade of 0. You **may not** copy code and/or answers **directly** from another student. If you copy someone else's work, both parties will receive a grade of 0.
* Rather than copying someone else's work, ask for help. You are not alone in this course!

---

## Course components:

- Teams: 3-4 person teams, based on survey and pretest results, consistent throughout the semester
- Application exercises: Usually start in class and finish in teams by the next class period, check/no check
- Homework: Individual, lowest score dropped
- Lab: Team, lowest score dropped
- Pyramid Exams: 3 parts, in class, take home, whole class 
- Final project: Presentations during the last week of class, you must participate in the project and be in class to present to pass this class
- Self paced tutorials: Individual, check/no check, extra credit

---

## Grading

Component             | Weight
----------------------|----------------
Participation & application exercises       | 10%
Peer evaluation       | 5%
Homework              | 20%
Labs                  | 15%
Midterm exam 1          | 15%
Midterm exam 2         | 15%
Final project         | 20%

- Class attendance is a firm expectation; frequent absences or tardiness 
will be considered a legitimate cause for grade reduction.

---

## Late/missed work policy

- Late work policy for homework assignments:
    - late, but within 24 hours of due date/time: -50%
    - any later: no credit
    
- Late work will not be accepted for take home midterms and the final project.

- Exam dates cannot be changed and no make-up exams will be given. If a midterm exam must be missed, absence must be officially excused in advance of the due date, in which case the missing exam score will be imputed using the other exam score. This policy only applies to the midterms.

- You must complete the final project and be in class to present it in order to pass this course.

---

## Other policies

- Please refrain from texting or using your computer for anything other than coursework during class.

- You must be in class on a day when you're scheduled to present, there are no make ups for presentations.

- Regrade requests must be made within 1 week of when the assignment is returned.

---
