---
title: "Simple linear regression"
author: ""
output:
  xaringan::moon_reader:
    css: "slides.css"
    lib_dir: libs
    nature:
      highlightLines: true
      highlightStyle: github
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```
layout: true

<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan
</span>
</div> 

---

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(emo)
library(dsbox)
```

class: center, middle

# Data = Model + Error

---

class: center, middle

$\Huge y = f(x) + \epsilon$

---

class: center, middle

$\Huge y = f(x) + \epsilon$

## Simple linear regression

---
class: center, middle

$\Huge \color{blue}y = f(x) + \epsilon$

* **y:** continuous (quantitative) variable

<br><br><br>

### properties of simple linear regression

---

class: center, middle

$\Huge y = f(\color{blue}x) + \epsilon$

* **x:** continuous (quantitative) variable

<br><br><br>

### properties of simple linear regression

---

class: center, middle

$\Huge y = \color{blue}{f(x)} + \epsilon$

* **f(x):** a function that gives the mean value of $y$ at any value of $x$

<br><br><br>

### properties of simple linear regression

---

.question[
What function do you think we are using get the mean value of $y$ with **simple linear regression**?
]
]

