---
title: "Prediction intervals"
author: ""
output:
  xaringan::moon_reader:
    css: "slides.css"
    lib_dir: libs
    nature:
      highlightLines: true
      highlightStyle: github
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```
layout: true

<div class="my-footer">
<span>
by Dr. Lucy D'Agostino McGowan
</span>
</div> 

---

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(Stat2Data)
data("Sparrows")
```

class: middle

# confidence intervals

If we use the same sampling method to select different samples and computed an interval estimate for each sample, we would expect the true population parameter ( $\beta_1$ ) to fall within the interval estimates 95% of the time.

---

# Confidence interval for $\beta_1$

.question[
How do we calculate the confidence interval for the slope?
]

--

.center[


$\Huge\hat\beta_1\pm t^*SE_{\hat\beta_1}$

]
---

## How do we calculate it in R?

* using the **broom** package

.small[
```{r, highlight.output = 5}
lm(Weight ~ WingLength, Sparrows) %>%
  tidy(conf.int = TRUE)
```
]

--

* "by hand"

.small[
```{r}
t_star <- qt(0.025, df = 116 - 2, lower.tail = FALSE)
# or
t_star <- qt(0.975, df = 116 - 2)
```

.pull-left[
```{r}
0.467 - t_star * 0.0347
```
]
.pull-right[
```{r}
0.467 + t_star * 0.0347
```
]
]

---

## Confidence intervals

There are `r emo::ji("v")` other types of confidence intervals we may want to calculate

--
* The confidence interval for the **mean response** in $y$ for a given $x^*$ value

--
* The confidence interval for an **individual response** $y$ for a given $x^*$ value

---

## Confidence intervals

There are `r emo::ji("v")` other types of confidence intervals we may want to calculate


* The confidence interval for the **mean response** in $y$ for a given $x^*$ value: **confidence interval for** $\mu_y$
* The confidence interval for an **individual response** $y$ for a given $x^*$ value: **prediction interval**

--
* Why are these different? Which do you think is easier to estimate?

--
* It is **harder** to predict one response than to predict a mean response. What does this mean in terms of the standard error?

--
* The SE of the prediction interval is going to be **larger**

---

## Confidence intervals

**confidence interval for** $\mu_y$ and **prediction interval**

.center[

$\Huge \hat{y}\pm t^* SE$

]

* $\hat{y}$ is the predicted $y$ for a given $x^*$
* $t^*$ is the critical value for the $t_{n-2}$ density curve
* $SE$ takes `r emo::ji("v")` different values depending on which interval you're interested in

--
* $SE_{\hat\mu}$

--
* $SE_{\hat{y}}$

--
* Which will be larger?

---

## Confidence intervals

**confidence interval for** $\mu_y$ and **prediction interval**

.center[

$\Huge \hat{y}\pm t^* SE$

]

* $\hat{y}$ is the predicted $y$ for a given $x^*$
* $t^*$ is the critical value for the $t_{n-2}$ density curve
* $SE$ takes `r emo::ji("v")` different values depending on which interval you're interested in
* $SE_{\hat\mu} = \hat{\sigma}_\epsilon\sqrt{\frac{1}{n}+\frac{(x^*-\bar{x})^2}{\Sigma(x-\bar{x})^2}}$
* $SE_{\hat{y}}=\hat{\sigma}_\epsilon\sqrt{1 + \frac{1}{n}+\frac{(x^*-\bar{x})^2}{\Sigma(x-\bar{x})^2}}$
* Which will be larger?

--
* What is the difference between these two equations?

---

## Confidence intervals

**confidence interval for** $\mu_y$ and **prediction interval**

.center[

$\Huge \hat{y}\pm t^* SE$

]

* $\hat{y}$ is the predicted $y$ for a given $x^*$
* $t^*$ is the critical value for the $t_{n-2}$ density curve
* $SE$ takes `r emo::ji("v")` different values depending on which interval you're interested in
* $SE_{\hat\mu} = \hat{\sigma}_\epsilon\sqrt{\frac{1}{n}+\frac{(x^*-\bar{x})^2}{\Sigma(x-\bar{x})^2}}$
* $SE_{\hat{y}}=\hat{\sigma}_\epsilon\sqrt{\color{red}1 + \frac{1}{n}+\frac{(x^*-\bar{x})^2}{\Sigma(x-\bar{x})^2}}$

--
* an individual response will vary from the mean response $\mu_y$ with a standard deviation of $\sigma_\epsilon$
---

## Let's do it in R!

.small[
```{r, eval = FALSE}
lm(Weight ~ WingLength, data = Sparrows) %>% 
  predict() 
```

```{r, echo = FALSE}
lm(Weight ~ WingLength, data = Sparrows) %>% 
  predict() %>% 
  head(3)
```
]

--

.small[
```{r, eval = FALSE}
lm(Weight ~ WingLength, data = Sparrows) %>% 
  predict(interval = "confidence")  #<<
```

```{r, echo = FALSE}
lm(Weight ~ WingLength, data = Sparrows) %>% 
  predict(interval = "confidence") %>% 
  head(3)
```
]

--

.small[
```{r, eval = FALSE}
lm(Weight ~ WingLength, data = Sparrows) %>%
  predict(interval = "prediction") #<<
```

`## WARNING predictions on current data refer to _future_ responses`
```{r, echo = FALSE, warning = FALSE}
lm(Weight ~ WingLength, data = Sparrows) %>%
  predict(interval = "prediction") %>% 
  head(3)
```
]

---

## Let's do it in R!

What if we have new data?
.small[
```{r}
new_sparrows <- data.frame(
  WingLength = c(30, 28, 25)
)
new_sparrows
```
]

--

.small[
```{r}
lm(Weight ~ WingLength, data = Sparrows) %>%
  predict(newdata = new_sparrows, interval = "prediction")
```

]
