---
title: "Adjusting and Interactions"
author: ""
output:
  xaringan::moon_reader:
    css: "slides.css"
    lib_dir: libs
    nature:
      highlightLines: true
      highlightStyle: github
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```
layout: true

<div class="my-footer">
<span>
by Dr. Lucy D'Agostino McGowan
</span>
</div> 

---

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(Stat2Data)
library(mosaicData)
knitr::opts_chunk$set(echo = FALSE)
```

## <i class="fas fa-laptop"></i> `Diamonds`

- Go to RStudio Cloud and open `Diamonds`

---

## Adjusting for confounders

* What is the relationship between average SAT scores and average teacher salaries?

.small[
```{r}
data("SAT")
DT::datatable(SAT, options = list(pageLength = 5))
```
]

--
* Are we doing inference or prediction?

---

## Adjusting for confounders

* I fit a linear model for $\hat{sat} = \hat\beta_0 + \hat\beta_1 salary$

```{r}
lm(sat ~ salary, SAT) %>%
  tidy()
```


--

* How do we interpret this result?

---

## Adjusting for confounders

* There is a **third variable**, the fraction of students that took the SAT in that state. It is grouped as "Low", "Medium", and, "High".

```{r}
SAT <- SAT %>%
  mutate(frac_group = case_when(
    frac < 22 ~ "LOW",
    frac < 49 ~ "MED",
    TRUE ~ "HIGH"
  ))
lm(sat ~ salary + frac_group, SAT) %>%
  tidy()
```

--

* What is the referent category?
--

* How do you interpret the $\hat{\beta}$ for `frac_groupLOW`?
--

* How do you interpret the $\hat{\beta}$ for `salary` now?

---

class: middle

## $\hat\beta$ interpretation in multiple linear regression

A one-unit increase in $x$ yields an expected increase in y of $\hat\beta$, **holding all other variables constant**.

---

class: middle

## $\hat\beta$ interpretation in multiple linear regression

A one-unit increase in average salary yields an expected increase in average SAT score of 1.09, **holding the fraction of students that took the SAT constant**.

---

## Adjusting for confounders

```{r}
ggplot(SAT, aes(salary, sat)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)
```

---

## Adjusting for confoundrs

```{r}
ggplot(SAT, aes(salary, sat, color = frac_group, group = frac_group)) + 
  geom_point() + 
  geom_line(aes(y = predict(lm(sat ~ salary + frac_group, data = SAT)))) +
  labs(color = "Fraction took SAT")
```

--

* What is this called? Where the direction reverses?
--

* Notice here the lines are **parallel** so holding the group constant, this is the effect we see.
--

* `r emo::ji("scream")` what if the lines aren't parallel?
---

## Interactions

* Data looking at the growth rate for kids

.small[
```{r}
data("Kids198")
DT::datatable(Kids198)
```
]

---

## Interactions

```{r}
ggplot(Kids198, aes(Age, Weight)) +
  geom_point()
```

--

* Will $\hat\beta_{age}$ be positive or negative?

---

## Interactions

* Let's look at this relationship split by `sex` (blue: Girl, black: Boy)

```{r}
ggplot(Kids198, aes(Age, Weight, color = Sex)) +
  geom_point() +
  theme(legend.position = "none")
```

---

## Interactions

* Let's look at this relationship split by `sex` (blue: Girl, black: Boy)

```{r}
ggplot(Kids198, aes(Age, Weight, color = Sex, group = Sex)) +
  geom_point() +
  theme(legend.position = "none") + 
  geom_smooth(method = "lm", se = FALSE)
```

--

* `r emo::ji("scream")` the lines cross! That means there is an **interaction**, that is the slopes differ based on the group

---

## Interactions

* Let's look at this relationship split by `sex` (blue: Girl, black: Boy)

```{r}
ggplot(Kids198, aes(Age, Weight, color = Sex, group = Sex)) +
  geom_point() +
  theme(legend.position = "none") + 
  geom_smooth(method = "lm", se = FALSE)
```

--

* What is the equation for this relationship?

---

## Interactions

$Weight = \beta_0 + \beta_1 Age + \beta_2 Girl + \beta_3 Age \times Girl + \epsilon$

.small[
```{r, echo = TRUE}
lm(Weight ~ Age + Sex + Age * Sex, data = Kids198)
```
]

--

* What does this model become for **boys** (When `Sex = 0`)

--
  * $Weight = \beta_0 + \beta_1 Age + \epsilon$

--
* What does this model become for **girls** (When `Sex = 1`)

--
  * $Weight = \beta_0 + \beta_1 Age + \beta_2 1 + \beta_3 Age \times 1 + \epsilon$

--
  * $Weight = (\beta_0 + \beta_2) + (\beta_1 + \beta_3) Age  + \epsilon$

--
* How do you interpret $\hat\beta_0$ now?

---

## Interactions

.small[
```{r, echo = TRUE}
lm(Weight ~ Age + Sex + Age * Sex, data = Kids198)
```
]


* What does this model become for **boys** (When `Sex = 0`)
  * $Weight = \beta_0 + \beta_1 Age + \epsilon$
* What does this model become for **girls** (When `Sex = 1`)
  * $Weight = \beta_0 + \beta_1 Age + \beta_2 1 + \beta_3 Age \times 1 + \epsilon$
  * $Weight = (\beta_0 + \beta_2) + (\beta_1 + \beta_3) Age  + \epsilon$
* How do you interpret $\hat\beta_{2}$ now?

--
  * The difference in intercepts between boys and girls

---

## Interactions

.small[
```{r, echo = TRUE}
lm(Weight ~ Age + Sex + Age * Sex, data = Kids198)
```
]


* What does this model become for **boys** (When `Sex = 0`)
  * $Weight = \beta_0 + \beta_1 Age + \epsilon$
* What does this model become for **girls** (When `Sex = 1`)
  * $Weight = \beta_0 + \beta_1 Age + \beta_2 1 + \beta_3 Age \times 1 + \epsilon$
  * $Weight = (\beta_0 + \beta_2) + (\beta_1 + \beta_3) Age  + \epsilon$
* How do you interpret $\hat\beta_{3}$ now?

--
  * How much the slope changes as we move from the regression line for boys to that for girls

---

## Interactions

$Weight = \beta_0 + \beta_1 Age + \beta_2 Girl + \beta_3 Age \times Girl + \epsilon$

* Hypothesis testing: What if you want to test whether the slope is different between groups?
* Is the growth rate different for boys and girls?
* What is $H_0$?

--
  * $H_0: \beta_3 = 0$

--
* What is $H_A$?

--
  * $H_A:\beta_3 \neq 0$
  
---

## Interactions

.small[
```{r, echo = TRUE, highlight.output = 7}
lm(Weight ~ Age + Sex + Age * Sex, data = Kids198) %>%
  tidy()
```
]

--

* What is the result of our hypothesis test?

---

## <i class="fas fa-laptop"></i> `Diamonds`

- Go to RStudio Cloud and open `Diamonds`

---


