---
title: "Variable Transformations Recap"
author: ""
output:
  xaringan::moon_reader:
    css: "slides.css"
    lib_dir: libs
    nature:
      highlightLines: true
      highlightStyle: github
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```
layout: true

<div class="my-footer">
<span>
by Dr. Lucy D'Agostino McGowan
</span>
</div> 

---

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(Stat2Data)
library(mosaicData)
knitr::opts_chunk$set(echo = FALSE)
```

class: middle

## $\hat\beta$ interpretation in multiple linear regression

The coefficient for $x$ is $\hat\beta$ (95% CI: $LB_\hat\beta, UB_\hat\beta$). A one-unit increase in $x$ yields an expected increase in y of $\hat\beta$, **holding all other variables constant**.

---

class: middle

## $\hat\beta_1$ interpretation in 

$sat = \beta_0 + \beta_1salary + \beta_2 (frac = LOW) + \beta_3 (frac = HIGH) + \epsilon$

The coefficient for average salary is 1.09 (95% CI: -0.90, 3.08).  A one-unit increase in average salary yields an expected increase in average SAT score of 1.09, **holding the fraction of students that took the SAT constant**.

---

## Adjusting for confoundrs

```{r}
SAT <- SAT %>%
  mutate(frac_group = case_when(
    frac < 22 ~ "LOW",
    frac < 51 ~ "MID",
    TRUE ~ "HIGH"
  ))
ggplot(SAT, aes(salary, sat, color = frac_group, group = frac_group)) + 
  geom_point() + 
  geom_line(aes(y = predict(lm(sat ~ salary + frac_group, data = SAT)))) +
  labs(color = "Fraction took SAT")
```

* The lines are **parallel**, the slope ( $\hat\beta_1$ ) is **constant** between groups


---

## Interactions


```{r}
data("Kids198")
ggplot(Kids198, aes(Age, Weight, color = Sex, group = Sex)) +
  geom_point() +
  theme(legend.position = "none") + 
  geom_smooth(method = "lm", se = FALSE)
```

--

* `r emo::ji("scream")` the lines cross! That means there is an **interaction**, that is the slopes differ based on the group

---

## Interactions

$Weight = \beta_0 + \beta_1 Age + \beta_2 Girl + \beta_3 Age \times Girl + \epsilon$

.small[
```{r, echo = TRUE}
lm(Weight ~ Age + Sex + Age * Sex, data = Kids198)
```
]

--

* What does this model become for **boys** (When `Sex = 0`)

--
  * $Weight = \beta_0 + \beta_1 Age + \epsilon$

--
* What does this model become for **girls** (When `Sex = 1`)

--
  * $Weight = \beta_0 + \beta_1 Age + \beta_2 1 + \beta_3 Age \times 1 + \epsilon$

--
  * $Weight = (\beta_0 + \beta_2) + (\beta_1 + \beta_3) Age  + \epsilon$

--
* How do you interpret $\hat\beta_0$ now?

---

## Interactions

.small[
```{r, echo = TRUE}
lm(Weight ~ Age + Sex + Age * Sex, data = Kids198)
```
]


* What does this model become for **boys** (When `Sex = 0`)
  * $Weight = \beta_0 + \beta_1 Age + \epsilon$
* What does this model become for **girls** (When `Sex = 1`)
  * $Weight = \beta_0 + \beta_1 Age + \beta_2 1 + \beta_3 Age \times 1 + \epsilon$
  * $Weight = (\beta_0 + \beta_2) + (\beta_1 + \beta_3) Age  + \epsilon$
* How do you interpret $\hat\beta_{2}$ now?

--
  * The difference in intercepts between boys and girls

---

## Interactions

.small[
```{r, echo = TRUE}
lm(Weight ~ Age + Sex + Age * Sex, data = Kids198)
```
]


* What does this model become for **boys** (When `Sex = 0`)
  * $Weight = \beta_0 + \beta_1 Age + \epsilon$
* What does this model become for **girls** (When `Sex = 1`)
  * $Weight = \beta_0 + \beta_1 Age + \beta_2 1 + \beta_3 Age \times 1 + \epsilon$
  * $Weight = (\beta_0 + \beta_2) + (\beta_1 + \beta_3) Age  + \epsilon$
* How do you interpret $\hat\beta_{3}$ now?

--
  * How much the slope changes as we move from the regression line for boys to that for girls

---

class: middle

## $\hat\beta$ interpretation for interactions between $x$ and a binary indicator $I$

The coefficient for the interaction between $x$ and $I$ is $\hat\beta$ (95% CI: $LB_\hat\beta, UB_\hat\beta$). This means that the effect of $x$ on $y$ differs by $\hat\beta$ when $I = 1$ compared to $I = 0$ **holding all other variables constant***.

--
* You must include this line if there are **additional variables in your model**.

---

class: middle

## $\hat\beta_3$ interpretation for  

$Weight = \beta_0 + \beta_1 Age + \beta_2 Girl + \beta_3 Age \times Girl + \epsilon$


The coefficient for the interaction between `Age` and `Sex` is -0.28 (95% CI: -0.44, -0.12). This means that the effect of `Age` on `Weight` lower by 0.28 among girls compared to boys.

---

## Non-linear relationships

```{r, echo = TRUE, eval = FALSE}
lm(TotalPrice ~ Carat + I(Carat^2), data = Diamonds)
```

```{r}
data("Diamonds")
ggplot(Diamonds, aes(Carat, TotalPrice)) +
  geom_point() + 
  geom_line(aes(y = predict(lm(TotalPrice ~ Carat + I(Carat^2), data = Diamonds))), lwd = 1.5, color = "blue")
```

--

* What is the equation for this relationship?

---

## Interpreting $\hat\beta$s in the presence of polynomials

$Total Price = \beta_0 + \beta_1 Carat + \beta_2 Carat^2 + \epsilon$

* What is the interpretation of $\hat\beta_1$?

--
* Typically, in multiple linear regression, the interpretation of $\hat\beta_i$ is:
a one-unit change in $x$ yields an expected change in $y$ of $\hat\beta_i$ **holding all other variables constant**.

--
  * What does it mean to see a change in `Caret` holding `Carat` $^2$ constant?

--
* When you have a polynomial term, you need to **specify the values you are changing between**, since the change is no longer constant across all values of $x$.

---

## Interpreting $\hat\beta$ in the presence of polynomials
.small[
```{r, echo = TRUE}
lm(TotalPrice ~ Carat + I(Carat^2), data = Diamonds) %>%
  tidy()
```
]

What is the expected change in `TotalPrice` for a one-unit change in `Carat`, changing from 0.8 to 1.8?

--

.pull-left[
.small[
```{r, echo = TRUE}
(-522.7 + 2386 * 1.8 + 4498.2 * 1.8^2) - 
  (-522.7 + 2386 * 0.8 + 4498.2 * 0.8^2)
```
]
]

--

.pull-right[
.small[
```{r, echo = TRUE}
2386 * (1.8 - 0.8) + 
  4498.2 * (1.8^2 - 0.8^2)
```
]
]

---

## Interpreting $\hat\beta$ in the presence of polynomials
.small[
```{r, echo = TRUE}
lm(TotalPrice ~ Carat + I(Carat^2), data = Diamonds) %>%
  tidy()
```
]

What is the expected change in `TotalPrice` for a one-unit change in `Carat`, changing from 1.8 to 2.8?

--

.small[
```{r, echo = TRUE}
2386 * (2.8 - 1.8) + 4498.2 * (2.8^2 - 1.8^2)
```
]

--
* Can we talk about $\hat\beta_1$ and $\hat\beta_2$ in the context of a one-unit change in `Carat`?

---
## Interpreting $\hat\beta$ in the presence of polynomials

* $\hat\beta$ coefficients that are transformations of the same $x$ variable **must** be interpreted together

--
* You must first choose to values of $x$ to change between, and then report the change. 

--
* A sensible choice for the two $x$ values can be the 25th% quantile and the 75th% quantile.

---

class: middle

## General $\hat\beta$ interpretation with quadratic terms

The linear term in the model for $x$ has a coefficient of $\hat\beta_1$ (95% CI: $(LB_{\hat\beta_1}, UB_{\hat\beta_1})$). The quadratic term in the model for $x$ has a coefficient of $\hat\beta_2$ (95% CI: $(LB_{\hat\beta_2}, UB_{\hat\beta_2})$). A change in $x$ from $a$ to $b$ yields an expected change in $y$ of $\hat\beta_1 (b - a) + \hat\beta_2 (b^2 - a^2)$ **holding all other variables constant***.

--
* You must include this line if there are **additional variables in your model**.

---

class: middle

## Specific $\hat\beta$ interpretation for $y = \beta_0 + \beta_1 Carat + \beta_2 Carat^2 + \epsilon$ model

The linear term in the model for `Carat` has a coefficient of 2386 (95% CI: $(906, 3866)$). The quadratic term in the model for `Carat` has a coefficient of $4498$ (95% CI: $(3981, 5016)$). A change in `Carat` from $0.7$ to $1.24$ yields an expected change in `TotalPrice` of $6000.5$.

--
* Where did I get 0.7 and 1.24?

---

## Quantiles

```{r, echo = TRUE}
Diamonds %>%
  summarise(q1 = quantile(Carat, 0.25),
            q3 = quantile(Carat, 0.75))
```

---

## <i class="fas fa-laptop"></i> `Diamonds`

- Go to RStudio Cloud and open `Diamonds`
- Fit the model  $TotalPrice = \beta_0  + \beta_1Carat + \beta_2 Carat^2 + \beta_3 Color+\epsilon$
- Find the 0.25 quantile and 0.75 quantile of `Carat`
- What is the interpretation of $\hat\beta_1$, $\hat\beta_2$, and $\hat\beta_3$?


