---
title: "Lab 06: Building reports in R"
author: ""
output:
  xaringan::moon_reader:
    css: "lab-slides.css"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

layout: true
  
<div class="my-footer">
<span>
Dr. Lucy D'Agostino McGowan
</span>
</div> 

---
```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(Stat2Data)
library(broom)
```

# Agenda

* **Knitting a pdf document**
* **Creating tables in R**
* **Lab 06:** on your own

---

## Knitting a pdf document

* Currently your yaml at the top of your `.Rmd` files looks something like this:

```yaml
---
title: "The title of your document"
name: "Your name"
output: html_document
---
```
---

## Knitting a pdf document

.question[
What do you think you would change to render this as a **pdf** document instead of **html**?
]

* Currently your yaml at the top of your `.Rmd` files looks something like this:

```yaml
---
title: "The title of your document"
name: "Your name"
output: html_document
---
```
---

## Knitting a pdf document

.question[
What do you think you would change to render this as a **pdf** document instead of **html**?
]

* Currently your yaml at the top of your `.Rmd` files looks something like this:

```{yaml}
---
title: "The title of your document"
name: "Your name"
output: pdf_document #<<
---
```

--

* For this lab and your final project I want you to create .pdf files.

---

## Making pretty tables

* We've been using the `tidy()` and `glance()` functions to print tables of our model output.
* These kind of look like "code" in our final reports
* We can make these prettier with the **knitr** package

```{r}
lm(wt ~ am, data = mtcars) %>%
  tidy()
```

---

## Making pretty tables

* We've been using the `tidy()` and `glance()` functions to print tables of our model output.
* These kind of look like "code" in our final reports
* We can make these prettier with the **knitr** package

--
* Load the **knitr** package with `library(knitr)`
* Use the `kable()` function to output a pretty table

```{r, eval = FALSE}
library(knitr) #<<
lm(wt ~ am, data = mtcars) %>%
  tidy() %>%
  kable() #<<
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(knitr) 
lm(wt ~ am, data = mtcars) %>%
  tidy() %>%
  kable(format = "html") 
```
---

## Putting all the R code at the end

* For these fancy reports, we want to see the R code, but we don't want it intersperced throughout the document. 
* You can "hide" an R chunk by adding `echo = FALSE` to each of your chunks.

````
`r ''````{r, echo = FALSE}
```
````
* You can "hide" all of your chunks by adding this to the top of your .Rmd file

```{eval = FALSE}
library(knitr)
opts_chunk$set(echo = FALSE)
```

--

* **THEN** at the end of you document (in the Appendix) you can add

````
`r ''````{r ref.label=knitr::all_labels(), echo = TRUE, eval = FALSE}
```
````

---

## Open `Lab 06: Building Reports` in RStudio

